---
title: "Alchoholics Data Analysis"
author: "Trisha, Lapi, Vaibava"
date: "11/14/2022"
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/Chart 3 - Distribution of students with amount of drinks consumed weekly.R")
source("../source/Chart 2 - Alcohol Consumption vs Life Expectancy.R")
source("../source/Chart 1 - Alcohol Consumption vs Happiness Score Scatterplot.R")
source("../source/summary_table.R")
source("../source/summary_info.R")
```

## Data Summary

Through the project we were able to figure out correlations through the different datasets and understand what they mean. Through the student alcohol data, we were able to see through the different variables that were calculated, such as **age**, **absences**, and **grades**, that student that drink more tend to do worse in their classes and are absent more.  Through the alcohol life expectancy, we can see, again through variables like beers, spirits, wine, and *mean* amount of alcohol (in liters), that there is a correlation between life expectancy and alcohol and the more people drink the lower their life expectancy gets. Finally, the happiness and alcohol consumption data told us that there is **little to no correlation** between the two, and that the two do not have an impact on each other.

## Chart 1
``` {r, echo = FALSE, warning = FALSE}
alcohol_vs_happiness_scatterplot
```

There's no easy way to say this, but there's absolutely **no** relation between happiness and alcohol consumption. With a scatter plot it's easiest to see; you can say that there's a slow rising trend in happiness, but considering there are many other factors such as *GDP per capita* and ability to actually buy alcohol there's absolutely no relation between alcohol consumption and happiness score. The happiest countries are those with the *most* money, and thereby the ability to afford more alcohol to consume.
life_expectancy_vs_alcohol_consumption
## Chart 2
``` {r, echo = FALSE, warning = FALSE}
life_expectancy_vs_alcohol_consumption
```

This chart once again shows, albeit inadvertently, a trend that we will come to hate. The fact that there are more factors than we can actually definitively study and draw a conclusion from that are affecting our perception of this chart. There are a variety of reasons why a country like Spain would have such high alcohol use while still having a high life expectancy, from **cultural reasons** as well as **population statistics** and even the quality of their healthcare. However, there is something important to note here: countries with incredibly high level of alcohol consumption, in excess of **11** liters, are not those with the highest life expectancy. Countries like Lithuania, Belarus, Grenada, these countries can have the best healthcare in the world and that still cannot fight the effects of alcoholism.

## Chart 3 

``` {r, echo = FALSE, warning = FALSE}
distribution_of_alcohol_use
```

This bar chart shows the distribution of students to how much they had to drink in any single week. This chart specifically shows that a common facet of cultural assumption, high schoolers drinking at an **increased** level compared to adults, is unfounded. This helps us to understand that the majority of students are not naturally alcoholics by circumstance of being students, and there are **other factors** that come into play when it comes to alcoholism rooting itself in students in high school.

## Table 

``` {r, echo = FALSE, warning = FALSE}
table_student()
```

``` {r, echo = FALSE, warning = FALSE}
table_life_expectancy()
```

``` {r, echo = FALSE, warning = FALSE}
table_happiness()
```

The three tables included above relate variables and analyze patterns regarding alcoholism within various countries. For example, we can see in the first table that there is a high correlation of individuals whos **Age, Studytime, Courses, and reputation** are affected by Alcohol Consumption. We also are able to notice that there is a high consumption with certain *middle eastern countries*. This is also backed by how there is a large amount of countries and states within the *northern hemisphere* who import large amounts of alcohol. 